---
title: "p8105_hw6_ha2546"
author: "Hana Akbarnejad"
date: "11/27/2019"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(rvest)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

## Load and clean data

```{r}

birthweight_data = read_csv("data/birthweight.csv") %>% 
  rename_at(vars(starts_with ("b")), 
  funs(str_replace(., "b", "baby_"))
  ) %>% 
  rename_at(vars(starts_with ("m")), 
  funs(str_replace(., "m", "mother_"))
  ) %>%
  rename_all(
  funs(str_replace(., "wt", "weight"))    
  ) %>% 
  rename(
    baby_sex = baby_abysex,
    father_race = frace,
    family_income = fincome,
    mother_delivery_weight = delweight,
    gestational_age = gaweeks,
    malformation = mother_alform,
    mother_menarche = mother_enarche,
    mother_delivery_age= mother_omage,
    past_live_birth = parity,
    past_low_birth = pnumlbw,
    past_small_birth = pnumsga,
    pre_preg_bmi = ppbmi,
    pre_preg_weight = ppweight,
    weight_gain = weightgain
  ) %>% 
  mutate(
    baby_sex = factor(baby_sex),
    father_race = factor(father_race),
    malformation = factor(malformation),
    mother_race = factor(mother_race)
  )

sum(is.na(birthweight_data))
```

factored the numeric
renamed variables


Firs, take a look at the distribution of birth weight:
```{r}

birthweight_data %>% 
  ggplot(aes(x = baby_weight)) + geom_density()
```
Although a little skewed to the left, the data is normally distributed.

Working on model:
To build this model, I use a combination of backward model building and the information I obtained through a brief research on factors affecting birthweight.
Variables I assume might be influential and I wan to start with: gestational age, mother delivery age, pre pregnancy BMI, mother's weight gain during pregnancy, number of cigarettes smoken during pregnancy, and family income.
```{r}

fit1 = birthweight_data %>% 
  lm(baby_weight ~ gestational_age + mother_delivery_age + pre_preg_bmi + weight_gain + smoken + family_income, data = .) 

fit1 %>% 
  broom::tidy()
```

explaining model:
##################

plotting the residuals against fitted values:
```{r}

birthweight_data %>% 
  add_residuals(fit1) %>% 
  add_predictions(fit1) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point() + geom_line(y = 0)
```

This is what we expect: residuals randomly distributed bouncing around zero.

# comparisons

first lets just see the models
```{r}

model1 = birthweight_data %>% 
  lm(baby_weight ~ baby_length + gestational_age, data = .)

model1 %>% 
  broom::tidy()

model2 = birthweight_data %>% 
  lm(baby_weight ~ baby_length + baby_head + baby_sex + baby_length*baby_head*baby_sex, data = .)

model2 %>% 
  broom::tidy()
```

Automated CV using modelr...
```{r}

# train/test split using resampling
cv_df = 
  crossv_mc(birthweight_data, 100) 

# conversting train and test lists to tibbles (I assume we don't have to because all lm???)
# cv_df = cv_df %>%
#   mutate(
#     train = map(train, as_tibble),
#     test = map(test, as_tibble)
#     )

# obtaining RMSE of our models:
cv_df = cv_df %>% 
  mutate(
    fit1  = map(train, ~lm(baby_weight ~ gestational_age + mother_delivery_age + pre_preg_bmi + weight_gain + smoken + family_income, data = .x)),
    model1  = map(train, ~lm(baby_weight ~ baby_length + gestational_age, data = .x)),
    model2  = map(train, ~lm(baby_weight ~ baby_length + baby_head + baby_sex + baby_length * baby_head * baby_sex, data = .x)) %>% 
  mutate(
    rmse_fit1 = map2_dbl(fit1, test, ~rmse(model = .x, data = .y)),
    rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y)))
  )

```

I used RMSE as a way to compare these models, and the plot below shows the distribution of RMSE values for each candidate model (fit1, model1, model2):

```{r}

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

